```{r echo=FALSE}
# Load packages
library(ggplot2)
library(dplyr)
library(corrplot)
library(PerformanceAnalytics)
library(psych)
library(Hmisc)
library(matrixcalc)
library(MASS)
library(gmodels)
```

```{r}
# Load data
data <- read.csv("https://raw.githubusercontent.com/miachen410/DATA605/master/train.csv")

# Data structure
str(data)
```

This dataset contains 1460 observations and 81 variables. The last variable SalePrice is the response variable (dependent variable) that we will be working with in the analysis below.

### Descriptive and Inferential Statistics. 

Provide univariate descriptive statistics and appropriate plots for the training data set.  Provide a scatterplot matrix for at least two of the independent variables and the dependent variable. Derive a correlation matrix for any three quantitative variables in the dataset.  Test the hypotheses that the correlations between each pairwise set of variables is 0 and provide an 80% confidence interval.  Discuss the meaning of your analysis.  Would you be worried about familywise error? Why or why not?

```{r}
# Univariate descriptive statistics
summary(data)
```

```{r}
# Focus on the summary statistics of SalesPrice
summary(data$SalePrice)
```

```{r}
# Distribution of SalePrice in a histogram

options(scipen = 5) # setting for not switching xticks to scientific notation

ggplot(data, aes(x = SalePrice)) +
  geom_histogram(fill="blue", binwidth = 10000) +
  scale_x_continuous(breaks = seq(0, 800000, by = 100000))
```


```{r}
# Choosing LotArea, OverallQual, GrLivArea and GarageArea with SalePrice as the correlation testing dataset
select_data <- data[, c("SalePrice", "LotArea", "OverallQual", "GrLivArea", "GarageArea")]
```

```{r}
# Scatterplot matrix
pairs.panels(select_data, method = "pearson") #correlation method
```


```{r}
# Correlation Matrix

corr_data <- rcorr(as.matrix(select_data, use = "complete.obs")) # only use observations that have complete data
corr_data # display correlation matrix

corr_matrix <- cor(select_data, use = "complete.obs")
corrplot(corr_matrix, type = "upper") # visualize the correlation matrix
```

We can see that the correlations are non-zero between the independent variables and the p-values are zero. Therefore, we can reject the null hypotheses that the correlations between each pairwise set of variables is 0. With that said, independent variables OverallQual, GrLivArea and GarageArea and LotArea each has a linear relationship with SalePrice, with OverallQual having the strongest correlation.


### Linear Algebra and Correlation.  

Invert your correlation matrix from above. (This is known as the precision matrix and contains variance inflation factors on the diagonal.) Multiply the correlation matrix by the precision matrix, and then multiply the precision matrix by the correlation matrix. Conduct LU decomposition on the matrix.  

```{r}
# Precision matrix
precision_matrix <- solve(corr_matrix)

round(precision_matrix, 2)
```

```{r}
# Multiply correlation matrix by precision matrix
round(corr_matrix %*% precision_matrix, 2)
```

```{r}
#Multiply precision matrix by correlation matrix
round(precision_matrix %*% corr_matrix, 2)
```

```{r}
# LU decomposition
Z <- lu.decomposition(corr_matrix)
Z

L <- Z$L
U <- Z$U
```

```{r}
# Test if L*U gives us the original correlation matrix
(L %*% U) == corr_matrix
```


### Calculus-Based Probability & Statistics.  

Many times, it makes sense to fit a closed form distribution to data.  Select a variable in the Kaggle.com training dataset that is skewed to the right, shift it so that the minimum value is absolutely above zero if necessary.  Then load the MASS package and run fitdistr to fit an exponential probability density function.  (See  https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/fitdistr.html ).  Find the optimal value of $\lambda$ for this distribution, and then take 1000 samples from this exponential distribution using this value (e.g., rexp(1000, $\lambda$)).  Plot a histogram and compare it with a histogram of your original variable.   Using the exponential pdf, find the 5th and 95th percentiles using the cumulative distribution function (CDF).   Also generate a 95% confidence interval from the empirical data, assuming normality.  Finally, provide the empirical 5th percentile and 95th percentile of the data.  Discuss.

```{r}
# Fit a variable to exponential distribution

hist(data$GrLivArea, breaks = 30) # GrLivArea is right-skewed

fit <- fitdistr(data$GrLivArea, "exponential") # Fit exponential distribution

lambda <- fit$estimate # Find optimal value of lambda
lambda
```

```{r}
# Take 1000 sample of the exponential distribution with lambda
set.seed(1234)
exp_lambda <- rexp(1000, lambda)

# Plot a histogram
ggplot(as.data.frame(exp_lambda), aes(exp_lambda)) +
  geom_histogram(binwidth = 300)
```

```{r}
# Find 5th and 95th percentiles of the exponential distribution
qexp(0.05, rate = lambda) # 5th percentile
qexp(0.95, rate = lambda) # 95th percentile
```

```{r warning=FALSE}
# Construct a 95% confidence interval from the empirical data, assuming normality
ci(data$GrLivArea, confidence = 0.95)
```

```{r}
# The empirical 5th and 95th percentiles
quantile(data$GrLivArea, c(0.05, 0.95))
```

The empirical 5th and 95th percentiles are very different from those from the fitted exponential distribution. This suggests that the exponential distribution is not a good fit for the data. 

The 95% confidence interval suggests that 95% of the time, we would expect to see a value between 1488.50 and 1542.44. However, this is built upon the mean. For this right-skewed variable, median would be a better representation of the data.


### Modeling.  

Build some type of multiple regression model and submit your model to the competition board.  Provide your complete model summary and results with analysis.  

```{r}
# Multiple Linear Regression Model
lm <- lm(SalePrice ~ OverallQual+GrLivArea+GarageArea+OverallQual:GrLivArea+OverallQual:GarageArea+GrLivArea:GarageArea, data = data)

summary(lm)
```

$R^2$ explains 78.19% of the variability and p-value is nearly zero, suggests that this relationship is not due to random variation.

Multiple Linear Equation:

$$
SalePrice = 82227.26 - 14415.61 \times OverallQual + 5.43 \times GrLivArea - 7.90 \times GarageArea \\
+ 14.49 \times OverallQual \times GrLivArea \\
+ 39.52 \times OverallQual \times GarageArea \\
- 0.1 \times GrLivArea \times GarageArea
$$

```{r}
# Residuals variability plot
plot(fitted(lm), resid(lm),
     xlab = "Sale Price", ylab = "Residuals",
     main = "Residuals of Sale Price")
abline(h = 0)
```

The residual plot seems to meet the constant variability condition with the residuals constantly above and below the zero line, with a few outliners.

```{r}
# Quantile-Quantile Plot
qqnorm(lm$residuals)
qqline(lm$residuals)
```

There is no significant curvature in the QQ plot; points tend to follow the straight line which suggests there is a linear relationship.
